---
title: "Simulation study"
format:
  html:
    toc: true
    df-print: kable
execute:
  echo: false
  warning: false
---

```{r}
#| include: false
library(dplyr)
library(tidyr)
library(knitr)
summ <- read.csv("renato-simulation-results.csv")
wide <- function(d) {
  summ |>
    filter(design == d) |>
    select(label, arm, rsq_trad_mean) |>
    pivot_wider(names_from = arm, values_from = rsq_trad_mean) |>
    rename(Scenario = label)
}
```

## Methods

- 100 x 70 lattice with KNN = 15
- Purely inductive out of smaple prediction
- Hold out sets have a completely independent graph
- 6 Fold CV:
    1. Block CV
    2. Random CV
- Simulated Autocorrelation:
    - X covariate with Moran's I of `0.379` and `0.686` respectively.
    - Errors at `0.390` and `0.770` respectively.
- Models used:
  - OLS
  - SEM (using GMM)
  - XGBoost
  - XGBoost + Spatial Lags
  - GraphSAGE
  - GrapheSAGE with node level LayerNorm
- Metric is $R^2 = 1 - \mathrm{SSE}/\mathrm{SST}$, averaged over six folds.

| Scenario | Mean function | Error |
|---|---|---|
| A | Linear | Independent |
| B | Nonlinear | Independent |
| C | Linear | Spatially autocorrelated |
| D | Nonlinear | Spatially autocorrelated |
| E | Linear with a spatial lag in the response | Independent |

## Results

Spatially blocked folds:

```{r}
kable(wide("block"), digits = 3)
```

Random folds:

```{r}
kable(wide("random"), digits = 3)
```

- Spatial Lags almost always improve the XGBoost model
- Layer normalization didn't help except at high levels of autocorrelation in the residuals

## Interpretation

- Nonlinearity (B, D) is where GraphSAGE Model improves the most
- Notably when there is a spatial lag in the response that where GraphSAGE works well as well (E).
- None of the DGPs are able to accurately observe the held out error so no amount of spatial lag in the covariates helps reconstructing.
