---
title: "Simulation study"
format:
  html:
    toc: true
    df-print: kable
execute:
  echo: false
  warning: false
---

```{r}
#| include: false
library(dplyr)
library(tidyr)
library(knitr)
summ <- read.csv("renato-simulation-results.csv")
wide <- function(d) {
  summ |>
    filter(design == d) |>
    select(label, arm, rsq_trad_mean) |>
    pivot_wider(names_from = arm, values_from = rsq_trad_mean) |>
    rename(Scenario = label)
}
```

## Methods

- Lattice 100 × 70, n = 7,000, neighbours from the 15 nearest nodes.
- Six inductive folds per scenario. Held-out nodes were given their own
  subgraph with no edges to training nodes.
- Folds built twice: spatially blocked and at random.
- Realized Moran's I: covariates 0.379 and 0.686, errors 0.390 and 0.770.
- Arms: least squares, spatial error model, gradient-boosted trees with and
  without spatially lagged covariates, GraphSAGE with and without per-node
  layer normalisation.
- Accuracy is $R^2 = 1 - \mathrm{SSE}/\mathrm{SST}$, averaged over six folds.

| Scenario | Mean function | Error |
|---|---|---|
| A | Linear | Independent |
| B | Nonlinear | Independent |
| C | Linear | Spatially autocorrelated |
| D | Nonlinear | Spatially autocorrelated |
| E | Linear with a spatial lag in the response | Independent |

## Results

Spatially blocked folds:

```{r}
kable(wide("block"), digits = 3)
```

Random folds:

```{r}
kable(wide("random"), digits = 3)
```

- GraphSAGE reached 0.954 to 0.968 on B and 0.928 to 0.937 on D, against
  0.734 to 0.778 and 0.386 to 0.456 for least squares.
- Covariate autocorrelation did not change that margin. On B, GraphSAGE fell
  from 0.968 to 0.954 as Ix rose from 0.0 to 0.7.
- On C, accuracy fell as Iu rose, and GraphSAGE fell fastest. At Iu = 0.8 it
  reached 0.006, against 0.280 for least squares and 0.285 for the spatial
  error model.
- E is the only process with a lag term. Lagged covariates raised
  gradient-boosted trees from 0.067 to 0.294 at Ix = 0.0 and from 0.231 to
  0.381 at Ix = 0.4. In A through D the same features moved accuracy by less
  than 0.05.
- Fold design altered the ranking. On E at Ix = 0.0, GraphSAGE reached 0.429
  blocked and 0.211 random. On C at Iu = 0.8, layer normalisation raised
  accuracy from 0.006 to 0.057 blocked and lowered it from 0.181 to 0.161
  random.
- Layer normalisation moved accuracy by less than the fold-to-fold standard
  deviation everywhere except C at Iu = 0.8.

## Interpretation

- The gain over parametric and tree-based arms was confined to a nonlinear
  mean (B, D) or a lag in the response (E).
- Where the mean was correctly specified and the residual field was
  spatially structured (C at Iu = 0.8), GraphSAGE was the worst arm tested.
- That follows from the protocol. No model observes the held-out error
  field, so a structured residual is not recoverable from neighbouring
  covariates, and added flexibility fits variation that does not transfer.
- Blocked folds match cross-region transfer. The two designs disagree by
  0.218 on E, so random-fold estimates do not carry over.

## Data

| File | Rows | Contents |
|---|---|---|
| `renato-scenario-folds-all.csv` | 792 | Every fold, scenario, arm, design |
| `renato-scenario-folds-block.csv` | 396 | Blocked folds |
| `renato-scenario-folds-random.csv` | 396 | Random folds |
| `renato-simulation-results.csv` | 132 | Means and standard deviations |

- Columns: `design`, `label`, `fold`, `arm`, `mae`, `rmse`, `rsq`,
  `rsq_trad`.
- `rsq` is squared Pearson correlation, invariant to affine rescaling.
  `rsq_trad` is $1 - \mathrm{SSE}/\mathrm{SST}$. They differ by up to 0.13
  where predictions are miscalibrated.
- All figures above are `rsq_trad`.
